Minimizing Time Required For Graduation

### 2010-12-11 paper/paper.tex
Added constraint to stop up from taking the same course twice in once quarter
fixed up the only take one course constraint to be in both LPs
initial rewrite of opening sections and diagram add

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{top=3cm,bottom=3cm,left=3cm,right=3cm,letterpaper} % or letterpaper (US) or a5paper or....

@@

% Adds a table of contents. Comment out if we don't want this.
\tableofcontents
\pagebreak

\abstract{The Bachelor's Degree is designed to take four years to complete;
however, many students fail to complete it this amount of time. One of the
factors contributing to this is a lack of proper planning during the
registration phase. In this paper we present an algorithm to assist students in
constructing a schedule that works towards completing a degree in the shortest
amount of time possible.}

\section{Introduction} \subsection{Background} Registration can be a stressful
time for every student at the University of Washington. Without deliberate
planning, students must decide which classes to take on a quarterly basis. The
decision process involves considering factors such as quarterly class
offerings, the amount of credits, the prerequisites of the classes, time
conflicts, and how the choices in current quarter affect future quarters. The
University of Washington registration system provides a schedule finder;
however, it only takes into account time conflicts, and does not give advice on
how to pick the most ideal schedule.

A student's graduation time may be extended because of his or her inability to
take a class that has limited offerings during that particular year.
Additionally, the advising office’s workload can also be increased due to the
changing plans and majors of students and their need for repeated advising.

\subsection{Goal} Our goal is to alleviate the problematic situation by creating
a mathematical algorithm that accomplishes the following goal: given a student's
current state and preferred credit hours per quarter, produce a class schedule
that minimizes the number of quarters until graduation, eliminates time
conflicts, and eliminates prerequisite conflicts.

\subsection{Importance} The implementation of the algorithm as part of the
school’s registration system will benefit the students by enabling them to map
out their quarterly class schedules until graduation. Students will be able to
preview their ideal schedule, the length of time until graduation, and the
credit hours per quarter.  This will lead to better decision making for what
majors to pursue, and given the student's current state, which schedules are the
most ideal in terms of saving time and resources. This will also help students
save money, as they will be in school for fewer quarters. The university also
benefits from the reduced labor cost and increased efficiency of the advising
offices.

\section{Problem Description} We focus on the classes offered within the Math
department, although our algorithm and model can be easily extended to include
over departments.

Every quarter the Math department offers a subset of classes from all of the
available courses, each class is offered during a certain subset of available
time periods. Since there are far more classes than time periods during the
regular five-day week there are inevitably time conflicts between courses that
a student could take. To graduate with a B.S. in Mathematics there are certain
classes that are required and then there are elective requirements too. Adding
to the problem is many classes cannot just be taken whenever they are available;
instead, certain prerequisites must be satisfied first. In the case of math
courses, there can be multiple different prerequisite sets that satisfy the
requirements for a class.

@@

\subsection{Definitions} We will now formalize some terms used in our paper.
\begin{itemize} \item A {\it course} refers to any course offered by the
University of Washington. A course has associated with it a name used in the
University of Washington time schedule, for example {\it MATH309}. A course also
has associated with it the quarters in which it is offered. Throughout this
paper, we will refer to these using their time schedule abbreviates: {\it F} for
Fall Quarter, {\it W} for Winter Quarter, {\it Sp} for Spring Quarter, and {\it
S} for Summer Quarter. Each course is also worth a specific amount of credits.
\item We refer to a specific instance of a course as a {\it course offering}.
Each course offering additionally has associated with it the time of the day at
which the course meets, the quarter it is being offered, and the section number
of the offering.  \end{itemize}

Additionally: \begin{itemize} \item Let $S$ be the set of courses required to be
taken in order to graduate. $S$ can be thought of as our goal state, and can
also contain elective courses a student simply wishes to take.  \item Let $T$ be
the set of courses a given student has already completed. If no courses have
been taken, this set may be empty.  \item Let $U$ be the set of all courses
a student could possibly ever take.  \item Let $P$ be a set of rules specifying
course prerequisites.  \item Let $l$ be the maximum number of credits a student
wishes to take during any given quarter.  \item Let $O$ be the set of all
50-minute time blocks between 8:30am and 7:30pm from Monday through Friday.
\end{itemize}

\subsection{Representing Prerequisites} Initially, we hoped to represent course
prerequisites using a directed acyclic graph. Each node was a course, and edges
existed between nodes $i$ and $j$ if and only if $i$ was a prerequisite for $j$.
However, it does not work in our model, and courses can have a complicated set
of multiple prerequisites.  For example, {\it MATH327} requires both $MATH126$
and $MATH300$, or just $MATH136$.

\begin{figure} [ht] 
\begin{center} 
\includegraphics[scale=0.4]{prereq_logic_example} 
\end{center} 
\caption{Evaluation of a rule} 
\label{logic_ex} \end{figure}

To solve this issues, we introduce a set of rules for each course. Each rule is
a boolean expression represented in propositional logic in disjunctive normal
form. For example, the rule associated with $MATH327$ in fig~\ref{logic_ex} is
$$ \text{{\it MATH327: (MATH126 AND MATH300) OR MATH136}}.  $$ As previously
stated, we denote the set of all rules for all classes as $P$.

Furthermore, to determine if a given student is eligible to register for
a course offering, we can evaluate the rule for the course, where each clause
takes on the value true if a student has already taken the course, and false if
he or she hasn't. If the rule evaluates to true, then the student is eligible to
take the course; otherwise, the student is note. For example, consider
figure~\ref{logic_ex}, here a student has previously taken courses $MATH126$ and
$MATH300$. Since $MATH126$ and $MATH300$ have been taken, then those literals in
the rule for $MATH327$ evaluate to true and therefore the entire rule is true.
Since the entire rule is true, the student can take $MATH327$ when next
available.

\subsection{Initial Model and Tractability} With a static schedule it would be
possible to deterministically plan an entire schedule from start to finish,
unfortunately we have neither the computational resources nor the luxury of
a static schedule to do plan. From the student’s point of view, although there
are some classes that are offered every quarter, senior electives are not, and
sometimes it cannot be predicted more than one quarter in advance when a senior
elective will be offered. Even given perfect information, where we know the
possible class schedules for every quarter we face another problem; finding the
best possible schedule reduces to enumerating possible class schedules for every
quarter until the graduation requirements are satisfied. Given that we cannot
predict in advance the conflicts between classes, we cannot even determine the
branching factor of the enumeration tree nor its depth. Since we can only
determine one quarter in advance, this suggests approaching the problem by using
heuristics to pick the best possible classes each quarter.

Instead of trying to solve the complete problem we chose to use the rolling
horizon approach used by Xiangtong et al \cite{xiangton:informs} when scheduling
classes for airlines. Our problem already has a well defined horizon
demarcation, the quarters. For the rest of the paper we focus on developing
methods to optimally pick classes in each quarter separately.

\section{Rolling Horizons Approach} We adapt a rolling horizons approach which
uses a heuristic to pick the best possible schedule for any given quarter, given
the classes offered during the quarter and the courses the student has
previously taken. We use a greedy algorithm that functions on a per-quarter
basis to accomplish this. The greedy algorithm greatly improves the tractability
of course scheduling over an initially unknown number of quarters.

@@

\subsection{Determining Available Course Offerings} We need to select the
classes we can currently take and then build conflict graphs for classes based
on the times they are offered. In this section, we let $L$ be the initially
empty set of course offerings a given student is eligible to take in the given
quarter, $q$.

\subsubsection{Procedure} We present the procedure for populating the set $L$ as
follows: \begin{enumerate} \item For each rule in $P$ we iterate over each
proposition in the rule, $p$, and check if $p \in T$. If $p \in T$ then we say
$p$ evaluates to true; otherwise, $p$ evaluates to false. We recall that each
rule is associated with a given course. If the entire rule evaluates to true,
then we add the course associated with it to $L$. Once this is complete, $L$
will consist of every course that could be taken based solely on prerequisites
that we have fulfilled.  \item We now prune this set by removing all courses
that are not offered in the current quarter. $L$ will necessarily now only
contain classes that we can take given the current quarter and the classes we
have already taken.  \item We now construct a conflict graph, $G_{conflict}$,
where every course offering in $L$ is a node, and there is an edge between two
nodes $i$ and $j$ if and only if the time of course offering $i$ overlaps the
time of course offering $j$.  \end{enumerate}

\subsubsection{Analysis} The selection procedure involves a lot of set queries
and iteration, but in asymptotic running time the algorithm is relatively quick.
Let $n$ be the total number of classes ever available to take, or in other
words, $n = \|U\|$. In the worst case, $P$ has a complex rule for all $n$
courses, and checking each rule in $P$ is linear in the number of propositions.
However, we know {\it a priori} from analysis of our data set that the largest
rule has twelve propositions, so we can bound the rule check time by a constant,
$C$, and therefore step one can be done in $O(Cn)$ time.

Continuing to step two, the worst case is when $L$ = $Q$. Checking whether
a course is offered in a given quarter can be done in constant time, and we have
$n$ courses to check, therefore step two can also be done in $O(Cn)$ time.

Finally, the last step is once again bounded by a constant and $n$. Considering
courses between 8:30am and 7:30pm, we have nine fifty-minute time periods, and
five days of the week, and therefore we have 45 time slots to check in our
conflict graph, again yielding $O(Cn)$ runtime.

@@

\subsection{Selection of Quarterly Classes} \subsubsection{Selection Procedure}
At this point in time, we know which course offerings are available for the
given student to register for, and we must now pick the most optimal ones.  We
present two ways of approaching quarterly class choice. We will initially encode
the class selection as a binary integer programming problem in spirit of the
methods presented by Dinkel et al \cite{dinkel:scheduling}. and Pritsker et al
\cite{prisker:informs}. A second approach developed by Yamada et al
\cite{yamada:heuristic} and Pferschy et al \cite{pferschy:kcg} is presented for
dealing with very large class choices where binary LP would be difficult solve.

\subsubsection{Heuristic Value Function} In order to determine which classes
should be picked in a given schedule, we use a heuristic function which assigns
each course offering a value. The higher the value of this function for a given
course offering is, the more likely the given course offering will be placed in
the schedule. Recalling that our goal is to minimize the total number of
quarters taken (and implicitly, maximize the total number of courses taken per
quarter). Our heuristic function should give higher value to course offerings
that will help us achieve this goal. One component to the heuristic function
then must be the number of dependent courses a given course offering has. If
a course satisfies many prerequisites, then we know that there are many courses
that we cannot complete without it, so we want our heuristic function to favor
it.

@@

@@

\begin{itemize}
    \item $ X_i = \left\{ \begin{array}{lr} 1 \text{ if we take } X_i\\ 0 \text{ if
we don't take } X_i \end{array} \right. $ 
    \item For all $t \in O$, we let $t$ represent the conflict graph of classes
    at time period $t$.
    \item Let $m = \|L\|$.  
\end{itemize} 
We now state our problem as follows: 
Maximize:

\begin{equation}
    \text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
    \label{qtr_val}
\end{equation}
Subject To:
\begin{equation}
    \sum_{i=1}^m Weight(c_i) * X_i \leq l, 
    \label{weight_lim}
\end{equation}

\begin{equation}
    \forall t \in O,X_i+X_j \leq 1, \forall pairs (X_i,X_j) \in t
    \label{time_con}
\end{equation}

@@

As noted in the reference literature this is actually a (strongly) NP-complete
problem as described \cite{pferschy:kcg}. For the rest of this section we
consider an alternative formulation of class selection as a knapsack problem
with conflict graphs and suggest the use a technique developed by Yamada et. al.
\cite{yamada:heuristic}.

We begin by simplifying $G_{conflict}$ to be a fully connected graph by adding
dummy nodes $v'$ and insert edges that connect $v'$ to every other node in the
graph. To ensure that $v'$ does not alter the final optimal solution, we assign
it the following properties: \begin{itemize} \item $Weight(v') = l+1$ \item
$Value(v') = 0$ \end{itemize} As the weight of our dummy node is greater than
our credit limit, and it's value is zero, we know that it will never appear in
the optimal solution. With the above reasoning we show for our case of class
schedule that Pferschy and Schauer's \cite{pferschy:kcg} claim that adding
a dummy node without loss of generality holds. We can now conveniently write the
class picking problem as the knapsack with conflict graphs problem (KCG):

Maximize:

\begin{equation}
    \text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
\end{equation}

Subject To:

\begin{equation}
\sum_{i=1}^m Weight(c_i) * X_i \leq l,
\end{equation}

\begin{equation}
X_i + X_j \leq 1 \text{ for all connected vertices } i, j \in G_{conflict},
\end{equation}

\begin{equation}
X_i \in \{0, 1\}
\end{equation}

For versions of this problem in the low hundreds of variables it is possible 
to solve this problem exactly using modern LP solving software 
\cite{yamada:heuristic}. In our case, just solving for the optimal classes for 
a single major fits well into the scope of a few hundred (or even just tens of) 
variables. For multiple majors or to consider class selection
for distribution and elective requirements, however the number of classes
balloons significantly and the KCG problem becomes much harder. For this version
of the problem we suggest using the techniques used by Yamada et al
\cite{yamada:heuristic}, which we briefly outline below.

>>>>

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{top=2cm,bottom=2cm,left=2cm,right=2cm,letterpaper} % or letterpaper (US) or a5paper or....

@@

\abstract{The Bachelor's Degree is designed to take four years to complete,
however, many students fail to complete it this amount of time. One 
factor contributing to this is poor planning when registering for classes.
We present an algorithm to assist students in
constructing a schedule that works towards completing a degree in the shortest
amount of time possible.}
\pagebreak

\section{Introduction} \subsection{Background} Registration can be a stressful
time for every student at the University of Washington. Without good
planning, students can make poor registration choices resulting in more quarter
being needed to graduate than otherwise necessary. Picking the correct class
involves considering quarterly class offerings, class credits, the prerequisites
of a class, time conflicts, and how to fulfill future prerequisites. The
University of Washington registration system already provides a tool to aid
students, schedule finder. Schedule finder however, only takes into account time
conflicts in one quarter, and does not consider what classes are best for
minimizing graduation time.

A student's graduation time may be extended because of his or her inability to
take a class that has limited offerings during that particular year.

\subsection{Goal} Our goal is to provide an automated system that gives quarter
by quarter schedules until a student graduates. As input our system will take in
the current set of classes a student has taken and the preferred credit hours
per quarter. Additional input will be the current and past university time
schedules. As output our system will provide a class schedule that minimizes the
number of quarters until graduation.

\subsection{Importance} The implementation of our system as part of the school’s
registration system will benefit the students by enabling them to map out their
quarterly class schedules until graduation. Students will be able to preview
schedules and the length of time until graduation.  This will lead to better
decision making for what majors to pursue, and given the student's current
state, which schedules are the most ideal in terms of saving time and resources.
This will also help students save money, as they will be in school for fewer
quarters. The university also benefits from the reduced labor cost and increased
efficiency of the advising offices.

\section{Problem Description} To make our examples concrete we will focus on the
classes offered within the Math department, although our algorithm and model can
be extended to include over departments.

Every quarter the Math department offers a subset of courses from all of the
available courses in the department. Each course is offered during a certain set
time periods. There are inevitably time conflicts between courses that a student
could take because there are far more courses offered than time periods during
the regular five-day week  To graduate with a B.S. in Mathematics there are
certain classes that are required and then there are elective requirements too.
Additionally many classes have prerequisites which much be satisfied before they
can be taken. In the case of math courses, there can be multiple, exclusive
prerequisite sets that independently satisfy the requirements for a class.

@@

\subsection{Definitions} We will now formalize some terms used in our paper.
\begin{itemize} \item A {\it course} refers to any course offered by the
University of Washington. A course is identified by its department name and
number, for example {\it MATH309}. Each course is also worth a specific amount
of credits. \item We refer to a specific instance of a course as a {\it course
offering}. Each course offering additionally has associated with it the time of
the day at which the course meets, the quarter it is being offered, and the
section number of the offering. Throughout this paper, we will refer to quarters
by their time schedule abbreviates: {\it F} for Fall Quarter, {\it W} for Winter
Quarter, {\it Sp} for Spring Quarter, and {\it S} for Summer Quarter.
\end{itemize}

Additionally: \begin{itemize} \item Let $S$ be the set of courses required to
graduate. $S$ can be thought of as our goal state, and can also contain elective
courses a student simply wishes to take.  \item Let $T$ be the set of courses
a given student has already completed. If no courses have been taken, this set
may be empty.  \item Let $U$ be the set of all courses a student could possibly
ever take.  \item Let $R$ be a set of rules specifying course prerequisites.
\item Let $l$ be the maximum number of credits a student wishes to take during
any given quarter.  \item Let $O$ be the set of all 50-minute time blocks
between 8:30am and 7:30pm from Monday through Friday. \item Let $L_i$ be the set
of courses a student can take in quarter i. \end{itemize}

\subsection{Representing Prerequisites} Initially, we hoped to represent course
prerequisites using a directed acyclic graph. Each node was a course, and edges
existed between nodes $i$ and $j$ if and only if $i$ was a prerequisite for $j$.
The DAG relationship works well when courses have single requirements. It
becomes difficult, however, to track whether requires are fulfilled with courses
that have multiple prerequisites, or worse courses that have alternative
prerequisites. We decided that it was would be feasible to model prerequisites
using a graph.
\begin{figure} [ht] \begin{center}
\includegraphics[scale=0.4]{prereq_logic_example} \end{center}
\caption{Evaluation of a rule} \label{logic_ex} \end{figure}

To represent alternative and multiple prerequisites we introduced a set of logic
rules for each course. The prerequisites for each course are expressed as
propositional logic statements stored in the set $R$. The literal propositions
in each rule are related to whether or not a student has taken a particular
course. For example, the rule associated with $MATH327$ in fig~\ref{logic_ex} is
$$ \text{{\it MATH327: (MATH126 AND MATH300) OR MATH136}}.$$

To determine if a given student is eligible to register for a course offering,
c we can evaluate the rule for c. Each literal proposition in the rule takes on
the value true if a student has already taken the course, and false if he or she
hasn't.  If the rule evaluates to true, then the student is eligible to take the
course; otherwise. For example, consider Figure~\ref{logic_ex}, here a student
has previously taken courses $MATH126$ and $MATH300$. Since $MATH126$ and
$MATH300$ have been taken, then those literals in the rule for $MATH327$
evaluate to true and therefore the entire rule is true.  Since the entire rule
is true, the student can take $MATH327$.

\subsection{Initial Model and Tractability} With a static schedule it would be
possible to plan an entire set of quarter by quarter schedules from start to
finish. Unfortunately we have neither the computational resources nor the luxury
of a static schedule. Some classes are offered every quarter others such as
senior electives are not. Furthermore for some classes it cannot be predicted
more than one quarter in advance when a senior elective will be offered. Even
given perfect information, where we know the possible class schedules for every
quarter we face another problem; finding the best possible schedule reduces to
enumerating possible class schedules for every quarter until the graduation
requirements are satisfied.  Given that we cannot predict in advance the
conflicts between classes, we cannot even determine the branching factor of the
enumeration tree nor its depth. Since we can only determine one quarter in
advance, this suggests approaching the problem by using heuristics to pick the
best possible classes each quarter.

Instead of trying to solve the complete problem we chose to use the "rolling
horizon approach" used by Xiangtong et al \cite{xiangton:informs} when
scheduling classes for airline pilots. Xiangtong et al. also could not develop
an optimal class schedule in one pass, instead they picked optimal classes based
using a heuristic over a well demarcated time period. Once they had a solution
to the smaller time period, they increase the size of the time period, or rolled
the horizon, and used the previously optimal solution to help determine
a solution to the new larger problem. The team's algorithm continued rolling the
time horizon on the problem until they had a solution to the entire problem.

\section{Picking Classes Per Quarter} Since determining an entire schedule at
once is intractable we adapted the rolling horizons approach which uses
a heuristic to pick the best possible schedule for a quarter. With a good
heuristic we will be able to chose classes that lead to a minimized graduation
time. Since we cannot analytically determine graduation time based on classes
chosen in a single quarter, our heuristic tries to give high value to classes
that would delay graduation is taken later rather than sooner.

@@

\subsection{Determining Available Course Offerings} We need to select the
classes we can currently take and then build conflict graphs for classes based
on the times they are offered. The set $L_i$ is built in this part of the
algorithm.

\subsubsection{Procedure} We present the procedure for populating the set $L$ as
follows: \begin{enumerate} \item For each rule $r \in R$ we iterate over each
proposition $p$ in the rule. Recall that $p$ represents a course, for example
$MATH125$. We then check if $p \in T$.  If $p \in T$ then we say $p$ evaluates
to true; otherwise, $p$ evaluates to false.  Using rules of logic we evaluate
$r$, if $r$ true, then we add the course associated with it to $L$.  Once this
is complete, $L$ will consist of every course that could be taken based solely
on prerequisites that we have fulfilled.  \item We now prune $L$ by removing all
courses that are not offered in the current quarter. $L$ now only contain
classes that we can take given the current quarter and the classes we have
already taken.  \item We now construct a conflict graph, $G_{conflict}$, where
every course offering in $L$ is a node, and there is an edge between two nodes
$i$ and $j$ if and only if the time of course offering $i$ overlaps the time of
course offering $j$.  \end{enumerate}

\begin{figure} [ht]
    \begin{center}
        \includegraphics[scale=0.35]{conflict_graph_ex}
    \end{center}
    \caption{A Conflict Graph}
    \label{conflictg}
\end{figure}

\subsubsection{Analysis} The selection procedure involves a lot of set queries
and iteration, but in asymptotic running time the algorithm is relatively quick.
Let $n$ be the total number of classes ever available to take, or in other
words, $n = \|U\|$. In the worst case, $P$ has a complex rule for all $n$
courses, and checking each rule in $P$ is linear in the number of propositions.
However, we know {\it a priori} from analysis of our data set that the largest
rule has twelve propositions, so we can bound the rule check time by a constant,
$C$, and therefore step one can be done in $O(n)$ time.

Continuing to step two, the worst case is when $L$ = {all classes offered during
the quarter}.  Checking whether a course is offered in a given quarter can be
done in constant time, and we have $n$ courses to check, therefore step two can
also be done in $O(n)$ time.

Finally, the last step is once again bounded by a constant and $n$. Considering
courses between 8:30am and 7:30pm, we have nine fifty-minute time periods, and
five days of the week, and therefore we have 45 time slots to check in our
conflict graph, again yielding $O(n)$ runtime.

@@

\subsection{Selection of Quarterly Classes} To select classes we give each class
a value and weight. The primary problem is the the classes contained in $L$
almost certainly have time conflicts, likely between courses that need to be
taken.

\subsubsection{Heuristic Value Function} In order to determine which classes
should be picked in a given schedule, we use a heuristic function which assigns
each class a value. The higher the value of this function for a class, the more
likely that class will be placed in the schedule.  Recalling that our goal is to
minimize the total number of quarters taken (and implicitly, maximize the total
number of courses taken per quarter).  Our heuristic function should give higher
value to class that maximize the number of classes taken this quarter and in
future quarters.  One component to the heuristic function then must be the
number of dependent courses a given course offering has. If a course satisfies
many prerequisites, then we know that there are many courses that we cannot
complete without it, so we want our heuristic function to favor it.

@@

\subsubsection{Selection Procedure} At this point in the algorithm we know which
courses are available to a student and their relative values. We present two
ways of approaching quarterly class choice.  We will initially encode the class
selection as a binary integer programming problem in spirit of the methods
presented by Dinkel et al \cite{dinkel:scheduling}. and Pritsker et al
\cite{prisker:informs}. A second approach developed by Yamada et al
\cite{yamada:heuristic} and Pferschy et al \cite{pferschy:kcg} is presented for
dealing with very large class choices where binary LP would be difficult solve.

@@

\begin{itemize}
    \item $ X_i = \left\{ \begin{array}{lr} 1 \text{ if we take } X_i\\ 0 \text{ if
we don't take } X_i \end{array} \right. $ 
    \item For all $t \in O$, we let $t$ represent the conflict graph of classes
    at time period $t$.
    \item Let $m = \|L\|$.  
\end{itemize} 
We now state our problem as follows:

Maximize: 
\begin{equation}
    \text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
    \label{qtr_val}
\end{equation}
Subject To:
\begin{equation}
    \sum_{i=1}^m Weight(c_i) * X_i \leq l, 
    \label{weight_lim}
\end{equation}

\begin{equation}
    \forall t \in O,X_i+X_j \leq 1, \forall \text{ connected pairs } (X_i,X_j) \in t
    \label{time_con}
\end{equation}

\begin{equation}
   X_i + X_j \leq 1, \forall \text{ pairs } (X_i,X_j) \text{ where } X_i \text{and} X_j \text{ are offerings of
   the same course}
    \label{course_con}
\end{equation}

@@

As noted in the reference literature this is actually a NP-hard
problem as described \cite{pferschy:kcg}. For versions of this problem in the
low hundreds of variables, however it is still possible to solve using modern LP
solving software \cite{yamada:heuristic}. In our case, just solving for the
optimal classes for a single major fits well into the scope of a few hundred (or
even just tens of) variables. For multiple majors or to consider class selection
for distribution and elective requirements, however the number of classes
balloons significantly and the KCG problem becomes much harder. For this version
of the problem we suggest the techniques used by Yamada et al
\cite{yamada:heuristic}, which we briefly outline below.

For the rest of this section we consider an alternative formulation of class
selection as a knapsack problem with conflict graphs and suggest the use
a technique developed by Yamada et. al.  \cite{yamada:heuristic}.
